Exploring mathematical objects from custom-tailored mathematical universes

### 2020-01-16 paper-filmat.tex
typo (thanks to Moritz L)

We argue that toposes and specifically the change in perspective provided by
toposes are ripe for philosophical analysis. In particular, there are the
following connections with topics in the philosophy of mathematics:
\begin{enumerate}
\item Toposes enrich the realism/anti-realism debate in that they paint the larger
picture that the platonic heaven of mathematical objects is not unique: besides
the standard heaven of the standard topos, we can fathom the alternate
heavens of all other toposes, all embedded in a secord-order heaven.
\item Mathematics is not only about studying mathematical objects, but also
about studying the relations between mathematical objects. The distinct view
on mathematical objects provided by any topos uncovers relations which
otherwise remain hidden.
% XXX: mention Olivia?
\item In some cases, a mathematical relation can be expressed quite succinctly
using the language of a specific topos and not so succinctly using the language
of the standard topos. This phenomenon showcases the importance of
\emph{appropriate language}.
\item Toposes provide new impetus to study constructive mathematics and
intuitionistic logic, in particular also to restrict to intuitionistic
logic on the meta level and to consider the idea that the platonic heaven might
be governed by intuitionistic logic.
% XXX contingent: To some extent, what the commonly agreed-upon rules of
% mathematics are is dependent on contingent historical factors. It is
% conceivable, for instance, that the intuitionists ...
\end{enumerate}
This note touches on these topics and invites further research.

>>>>

We argue that toposes and specifically the change in perspective provided by
toposes are ripe for philosophical analysis. In particular, there are the
following connections with topics in the philosophy of mathematics:
\begin{enumerate}
\item Toposes enrich the realism/anti-realism debate in that they paint the larger
picture that the platonic heaven of mathematical objects is not unique: besides
the standard heaven of the standard topos, we can fathom the alternate
heavens of all other toposes, all embedded in a second-order heaven.
\item Mathematics is not only about studying mathematical objects, but also
about studying the relations between mathematical objects. The distinct view
on mathematical objects provided by any topos uncovers relations which
otherwise remain hidden.
% XXX: mention Olivia?
\item In some cases, a mathematical relation can be expressed quite succinctly
using the language of a specific topos and not so succinctly using the language
of the standard topos. This phenomenon showcases the importance of
\emph{appropriate language}.
\item Toposes provide new impetus to study constructive mathematics and
intuitionistic logic, in particular also to restrict to intuitionistic
logic on the meta level and to consider the idea that the platonic heaven might
be governed by intuitionistic logic.
% XXX contingent: To some extent, what the commonly agreed-upon rules of
% mathematics are is dependent on contingent historical factors. It is
% conceivable, for instance, that the intuitionists ...
\end{enumerate}
This note touches on these topics and invites further research.
